---
title: "30DayChartChallenge 2022-04-01"
description: |
  A short description of the post.
author:
  - name: Charlie Joey Hadley
    url: {}
date: 2022-03-31
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
set.seed(as.integer(Sys.Date()))
sample(1:5, 1)
sample(1:52, 1)
```

Here's the gist https://gist.github.com/charliejhadley/0a1375ac8f93d480550b61b4806aaa0a
and the tweet https://twitter.com/charliejhadley/status/1509893261875425297


Today's #30DayChartChallenge is "part-to-whole" and I decided to use it as an opportunity to make a waffle chart as I've never done it before. I'm going to try not to use the fivethirtyeight::bechdel dataset for every challenge, but I do really like it.

So let's start by counting how many films fit into each test category, and tidy these up

```{r }
library(fivethirtyeight)
library(tidyverse)
bechdel_count_clean_test <- bechdel %>% 
  count(clean_test)

label_tests_vec <- c("nowomen" = "No women in the movie", "notalk" = "No talking between women", "men" = "Talk about men", "dubious" = "Dubiously passes", "ok" = "Passed the test!")
label_tests_tib <- enframe(label_tests_vec) %>% 
  rename(clean_test = name,
         test_label = value)


bechdel_count_clean_test <- bechdel %>% 
  count(clean_test, sort = TRUE) %>% 
  left_join(label_tests_tib) %>% 
  mutate(test_label = fct_reorder(test_label, n))

bechdel_count_clean_test %>% 
  knitr::kable()
```

Looks to me like the [{waffle}](https://github.com/hrbrmstr/waffle) package is the best for making waffle charts with {ggplot2} but there are alternatives too. 

It was fun when making this to explore with different shapes of the chart. I initially played around with prime factors of `nrow(bechdel)` but eventually settled on the most square possible chart via `round(sqrt(nrow(bechdel)))`.

```{r}
library(waffle)
library(ggpomological)
library(hrbrthemes)
library(ggtext)
gg_waffle_bechdel <- bechdel_count_clean_test %>% 
  ggplot(aes(fill = test_label, values = n)) +
  geom_waffle(n_rows = round(sqrt(nrow(bechdel))), size = 0.33, colour = "white", flip = TRUE) +
  coord_equal()
gg_waffle_bechdel
```

Let's beautify:

```{r}
bechdel_count_clean_test %>% 
  ggplot(aes(fill = test_label, values = n)) +
  geom_waffle(n_rows = round(sqrt(nrow(bechdel))), size = 0.33, colour = "white", flip = TRUE) +
  scale_fill_pomological(name = "") +
  coord_equal() +
  theme_ipsum_rc(grid="") +
  theme_enhance_waffle()
```

Let's do this by decade. Again, it was fun to experiment with different `n_rows()` values. It feels to me like the best option is to use prime factors of the decade with most observations, which became 7 * 2.

```{r}
bechdel_by_decade <- bechdel %>% 
  mutate(decade = floor(year / 10) * 10) %>% 
  count(decade, clean_test, sort = TRUE) %>% 
  left_join(label_tests_tib) %>% 
  mutate(test_label = fct_reorder(test_label, n))

bechdel_by_decade %>% 
  ggplot(aes(fill = clean_test, values = n)) +
  geom_waffle(n_rows = 7 * 2, size = 0.33, colour = "white", flip = TRUE) +
  facet_wrap(~decade, nrow = 1, strip.position = "bottom") +
  scale_fill_pomological(name = "") +
  coord_equal() +
  theme_ipsum_rc(grid="") +
  theme_enhance_waffle()
```

Now I also haven't really experimented with patchwork


```{r}
library(cowplot)

gg_bechdel_waffle <- bechdel_count_clean_test %>% 
  ggplot(aes(fill = test_label, values = n)) +
  geom_waffle(n_rows = round(sqrt(nrow(bechdel))), size = 0.33, colour = "white", flip = TRUE, show.legend = TRUE) +
  scale_fill_pomological(name = "") +
  guides(fill = guide_legend(nrow = 2)) +
  coord_equal() +
  theme_ipsum_rc(grid="") +
  theme_enhance_waffle() +
  theme(legend.position = "bottom")
```


```{r}
gg_bechdel_decade_prop <- bechdel_by_decade %>% 
  mutate(test_label = fct_relevel(test_label, label_tests_tib$test_label)) %>% 
  ggplot(aes(fill = test_label, values = n)) +
  geom_waffle(n_rows = 10, size = 0.33, colour = "white", flip = TRUE, show.legend = TRUE, make_proportional = TRUE) +
  facet_wrap(~decade, nrow = 2, strip.position = "bottom") +
  scale_fill_ipsum(name = "") +
  labs(subtitle = paste("{waffle} charts are pretty fun!",
                      "<br>",
                      "<b>Square waffles</b>",
                      "<br>",
                      str_wrap("For the squarest waffles use this code: round(sqrt(nrow(data))). It'd be fun if this was an optional argument for geom_waffle()", 20),
                      "<br>",
                      "<b>Proportional Faceted waffles</b>",
                      "<br>",
                      "With proportional faceting n_rows should be a factor of 10. If not proportional, it feels like the best option is finding prime factors for the largest group. Consider using numbers::primeFactors() for that",
                      "<br>",
                      "<b>Paragraphs of text</b>",
                      "<br>",
                      "I've not really used {patchwork} or {cowplot} to add paragraphs of text. I'm always really impressed by folks who do have these nicely displayed summaries in their charts. So that's something I'm going to experiment with a lot this edition of #30DaysChartChallenge",
                      "<br>"
  )) +
  coord_equal() +
  theme_ipsum_rc(grid="") +
  theme_enhance_waffle() +
  theme(
    # plot.subtitle = element_markdown(lineheight = 1.25, hjust = 0, size = 13),
        plot.subtitle = element_textbox_simple(
      size = 10,
      padding = margin(2, 2, 2, 2),
      margin = margin(0, 0, 2, 0),
      lineheight = 1.25
    ),
        text = element_text(colour = "#242c28", family = "Arvo"))

```

```{r}
cw_bechdel_waffles <- plot_grid(
  ggdraw() + 
  draw_label(
    "#30DayChartChallenge 2022-04-01 Part to Whole: Waffles and the Bechdel Test",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, -10, 10),
    plot.title = element_text(family = "Arvo")
  ),
  plot_grid(gg_bechdel_waffle,
  gg_bechdel_decade_prop,
  nrow = 1,
  rel_widths = c(2, 3)),
  ncol= 1,
  rel_heights = c(0.05, 1)
)

cw_bechdel_waffles %>% 
  ggsave(here::here("_posts", "2022-03-31-30daychartchallenge-2022-04-01", "cw_bechdel_waffles.png"),
         .,
         width = 18,
         height = 8)
```


```{r}
cw_bechdel_waffles <- plot_grid(
  gg_bechdel_waffle,
  plot_grid(grid::textGrob(str_glue("{{waffle}} charts are pretty fun.",
                                    "\n",
                                    "For the squarest waffles use `round(sqrt(nrow(data)))``. It'd be fun if this was an optional argument for geom_waffle()",
                                    "\n",
                                    "Automatic pretty breaks for facetted geom_waffle() is a little harder. It feels like the best option is finding prime factors for the largest group."
                                    ),
                           just = "left",
                           x = unit(0, "npc")),
            gg_bechdel_decade_waffle,
            nrow = 2,
            rel_widths = c(1, 1)),
  ncol = 2
)
```



```{r}
plot_grid(
  gg_bechdel_waffle,
  plot_grid(ggdraw() +
  draw_label(str_glue("{{waffle}} charts are pretty fun.",
                      "\n",
                      "For the squarest waffles use `round(sqrt(nrow(data)))``. It'd be fun if this was an optional argument for geom_waffle()",
                      "\n",
                      "Automatic pretty breaks for facetted geom_waffle() is a little harder. It feels like the best option is finding prime factors for the largest group."
  ), hjust = 0),
            gg_bechdel_decade_waffle,
            nrow = 2),
  ncol = 2
)
```




```{r}
plot_grid(
  gg_bechdel_waffle,
  plot_grid(gg_bechdel_decade_waffle,
            nrow = 2),
  ncol = 2
)
```



```{r}
cw_bechdel_waffles %>% 
  ggsave(here::here("_posts", "2022-03-31-30daychartchallenge-2022-04-01", "cw_bechdel_waffles.png"),
         .,
         width = 18,
         height = 12)
```











